---
title: "policy_data"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{policy_data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message = FALSE}
library(polle)
```

This vignette is a guide to `policy_data()`. As the name suggests, the function creates a `policy_data` object with
a specific data structure making it easy to use in combination with `policy_def()`, `policy_learn()`, and `policy_eval()`.
The vignette is also a guide to some of the associated S3 functions
which transform or access parts of the data, see `?policy_data` and `methods(class="policy_data")`.

We will start by looking at a simple single-stage example, then consider a fixed two-stage example with varying actions sets and data in wide format, and
finally we will look at an example with a stochastic number of stages and data in long format.

# Single-stage: wide data

Consider a simple single-stage problem with covariates/state variables $(Z, L, B)$, binary action variable $A$, and
utility outcome $U$. We use `sim_single_stage()` to simulate data:

```{r single stage data}
(d <- sim_single_stage(n = 5e2, seed=1)) |> head()
```

We give instructions to `policy_data()` which variables define the `action`, the state `covariates`, and the `utility` variable: 

```{r pdss}
pd <- policy_data(d, action="A", covariates=list("Z", "B", "L"), utility="U")
pd
```

In the single-stage case the history $H$ is just $(B, Z, L)$. We access the history and actions using
`get_history()`:

```{r gethistoryss}
get_history(pd)$H |> head()
get_history(pd)$A |> head()
```

Similarly, we  access the utility outcomes $U$:

```{r get}
get_utility(pd) |> head()
```

```{r cleanup, include=FALSE}
rm(list = ls())
```

# Two-stage: wide data

Consider a two-stage problem with observations $O = (B, BB, L_{1}, C_{1}, U_{1}, A_1, L_2, C_{2}, U_{2}, A_2, U_{3})$. Following the general notation introduced in Section 3.1 of the associated paper, $(B,BB)$ are the baseline covariates, $S_k =(L_{k, C_{k}})$ are the
state covariates at stage k, $A_{k}$ is the action at stage k, and $U_k$ is the reward at stage $k$.
The utility is the sum of the rewards $U=U_{1}+U_{2}+U_{3}$.

We use `sim_two_stage_multi_actions()` to simulate data:
```{r simtwostage}
d <- sim_two_stage_multi_actions(n=2e3, seed = 1)
colnames(d)
```
Note that the data is in wide format.
The data is transformed using `policy_data()` with instructions on which
variables define the actions, baseline covariates, state covariates, and the rewards:

```{r pdtwostage}
pd <- policy_data(d,
                  action = c("A_1", "A_2"),
                  baseline = c("B", "BB"),
                  covariates = list(L = c("L_1", "L_2"),
                                    C = c("C_1", "C_2")),
                  utility = c("U_1", "U_2", "U_3"))
pd
```

The length of the character vector `action` determines the number of stages `K` (in this case 2).
If the number of stages is 2 or more, the `covariates` argument must be a named list. Each element must be
a character vector with length equal to the number of stages. If a covariate is not available at a
given stage we insert an `NA` value, e.g., `L = c(NA, "L_2")`.

Finally, the `utility` argument must
be a single character string (the utility is observed after stage K) or a character vector
of length K+1 with the names of the rewards.

In this example, the observed action sets vary for each stage. `get_action_set()` returns the
global action set and `get_stage_action_sets()` returns the action set for each stage:

```{r getactionsets}
get_action_set(pd)
get_stage_action_sets(pd)
```

The full histories $H_1 = (B, BB, L_{1}, C_{1})$ and $H_2=(B, BB, L_{1}, C_{1}, A_{1}, L_{2}, C_{2})$ are available using `get_history()` and `full_history = TRUE`:

```{r gethistwostage}
get_history(pd, stage = 1, full_history = TRUE)$H |> head()
get_history(pd, stage = 2, full_history = TRUE)$H |> head()
```
Similarly, we access the associated actions at each stage via list element `A`:

```{r}
get_history(pd, stage = 1, full_history = TRUE)$A |> head()
get_history(pd, stage = 2, full_history = TRUE)$A |> head()
```

Alternatively, the state/Markov type history and actions are available using `full_history = FALSE`:

```{r gethisstate}
get_history(pd, full_history = FALSE)$H |> head()
get_history(pd, full_history = FALSE)$A |> head()
```

Note that `policy_data()` overrides the action variable names to `A_1`, `A_2`, ... in the full history case and
`A` in the state/Markov history case.

As in the single-stage case we access the utility, i.e. the sum of the rewards, using
`get_utility()`:

```{r getutiltwo}
get_utility(pd) |> head()
```

# Multi-stage: long data

In this example we illustrate how `polle` handles decision
processes with a stochastic number of stages, see Section 3.5 of the associated paper.
The data is simulated using `sim_multi_stage()`.
Detailed information on the simulation is available in `?sim_multi_stage`.
We simulate data from 2000 iid subjects:

```{r sim_data}
d <- sim_multi_stage(2e3, seed = 1)
```
As described, the stage data is in long format:

```{r view_data}
d$stage_data[, -(9:10)] |> head()
```

The `id` variable is important for identifying which rows belong
to each subjects. The baseline data uses the same `id` variable:

```{r view_b_data}
d$baseline_data |> head()
```

The data is transformed using `policy_data()` with `type = "long"`.
The names of the `id`, `stage`, `event`, `action`,
and `utility` variables must be specified. The event variable, inspired by
the event variable in `survival::Surv()`, is `0` whenever an
action occur and `1` for a terminal event.

```{r pd}
pd <- policy_data(data = d$stage_data,
                  baseline_data = d$baseline_data,
                  type = "long",
                  id = "id",
                  stage = "stage",
                  event = "event",
                  action = "A",
                  utility = "U")
pd
```

In some cases we are only interested in analyzing a subset of the decision stages.
`partial()` trims the maximum number of decision stages:

```{r partial}
pd3 <- partial(pd, K = 3)
pd3
```
